Apache Flink\texttrademark : Streaming Dataflows as a Basis \\ for Universal Data Analytics

### 2015-11-10 intro.tex
Links moved to footnote

It is becoming more and more apparent, however, that a huge number of today's large scale data processing use cases handle data that is, in reality, continuously produced over time. These continuous streams of data come for example from web logs, application logs, sensors, or as changes to application state in databases (transaction log records). Rather than treating the streams as streams, today's setups ignore the continuous and timely nature of the data production. Instead, data records are (often artificially) batched into static data sets (for example hourly/daily/monthly chunks) and then processed in a time-agnostic fashion. Data collection tools, workflow managers, and schedulers orchestrate the creation and processing of batches, in what is actually a continuous data processing pipeline. Architectural patterns like the "lambda architecture" [] combine batch and stream processing systems to implement multiple paths of computation: A streaming fast path for timely approximate results, and a batch offline path for late accurate results.

Apache Flink follows a paradigm that embraces data stream processing as the unifying model for real-time analysis, continuous streams, and batch processing, both in the programming model and in the execution engine. In combination with durable message queues that allow quasi arbitrary replay of data streams (like Apache Kafka [] or Amazon Kinesis []), stream processing programs make no difference between processing the latest events in real-time, continuously aggregating data periodically in large windows, or processing terabytes of historical data: these types of computations simply start their processing at different points in the durable stream, and maintain different forms of state during the computation. Through a highly flexible windowing mechanism, Flink programs can compute both early and approximate, as well as delayed and accurate results in the same operation, obviating the need to combine different systems for the two use cases. Flink supports different notions of time (event time, ingestion time, processing time) in order to give programmer high flexibility in defining how events should be correlated.

At the same time, Flink acknowledges that there is, and will be, a need for dedicated batch processing (dealing with static data sets). Complex queries over static data are still a good match for a batch processing abstraction. Furthermore, batch processing is still needed both for legacy implementations of streaming use cases, and for analysis applications where no efficient algorithms are yet known that perform this kind of processing on streaming data. Batch programs are special cases of streaming programs, where the stream is finite, and order/time of records does not matter (all records implicitly belong to one all-encompassing window). However, to support batch use cases with a competitive ease and performance, Flink has a specialized API for processing static data sets, uses specialized data structures and algorithms for the batch versions of operators like join or grouping, and uses dedicated scheduling strategies. The result is that Flink presents itself as a full-fledged and efficient batch processor on top of a streaming runtime, including libraries for graph analysis and machine learning. 
Flink is a top-level project of the Apache Software Foundation that is developed and supported by a large and lively community (of more than 140 open source contributors at the time of this writing []), and is used in production in several companies.

>>>>

It is becoming more and more apparent, however, that a huge number of today's large scale data processing use cases handle data that is, in reality, continuously produced over time. These continuous streams of data come for example from web logs, application logs, sensors, or as changes to application state in databases (transaction log records). Rather than treating the streams as streams, today's setups ignore the continuous and timely nature of the data production. Instead, data records are (often artificially) batched into static data sets (for example hourly/daily/monthly chunks) and then processed in a time-agnostic fashion. Data collection tools, workflow managers, and schedulers orchestrate the creation and processing of batches, in what is actually a continuous data processing pipeline. Architectural patterns like the "lambda architecture"~\cite{marz2015big} combine batch and stream processing systems to implement multiple paths of computation: A streaming fast path for timely approximate results, and a batch offline path for late accurate results.

Apache Flink follows a paradigm that embraces data stream processing as the unifying model for real-time analysis, continuous streams, and batch processing, both in the programming model and in the execution engine. In combination with durable message queues that allow quasi arbitrary replay of data streams (like Apache Kafka\footnote{http://kafka.apache.org/} or Amazon Kinesis\footnote{https://aws.amazon.com/kinesis/}), stream processing programs make no difference between processing the latest events in real-time, continuously aggregating data periodically in large windows, or processing terabytes of historical data: these types of computations simply start their processing at different points in the durable stream, and maintain different forms of state during the computation. Through a highly flexible windowing mechanism, Flink programs can compute both early and approximate, as well as delayed and accurate results in the same operation, obviating the need to combine different systems for the two use cases. Flink supports different notions of time (event time, ingestion time, processing time) in order to give programmer high flexibility in defining how events should be correlated.

At the same time, Flink acknowledges that there is, and will be, a need for dedicated batch processing (dealing with static data sets). Complex queries over static data are still a good match for a batch processing abstraction. Furthermore, batch processing is still needed both for legacy implementations of streaming use cases, and for analysis applications where no efficient algorithms are yet known that perform this kind of processing on streaming data. Batch programs are special cases of streaming programs, where the stream is finite, and order/time of records does not matter (all records implicitly belong to one all-encompassing window). However, to support batch use cases with a competitive ease and performance, Flink has a specialized API for processing static data sets, uses specialized data structures and algorithms for the batch versions of operators like join or grouping, and uses dedicated scheduling strategies. The result is that Flink presents itself as a full-fledged and efficient batch processor on top of a streaming runtime, including libraries for graph analysis and machine learning. 
Flink is a top-level project of the Apache Software Foundation that is developed and supported by a large and lively community (of more than 140 open source contributors at the time of this writing~\cite{CUSTOM:web/FlinkContrib}), and is used in production in several companies.

### 2015-11-10 related.tex
Links moved to footnote

\para{Batch Processing.} Apache Hadoop~\cite{CUSTOM:web/Hadoop} is the most popular open source system for large-scale data analysis that is based on the MapReduce paradigm~\cite{DBLP:journals/cacm/DeanG08}. Dryad~\cite{isard2007dryad}, introduced embedded user-defined functions in general DAG-based dataflows and was used by SCOPE~\cite{scopeOptimizer} which added a language and an SQL optimizer on top of it. Apache Tez~\cite{CUSTOM:web/Tez} can be seen as an open source implementation of the ideas proposed in Dryad. MPP databases \cite{dewitt1990gamma} , and recent open-source implementations~\cite{CUSTOM:web/Drill,kornacker2015impala} restrict their API to SQL variants. Similar to Flink, Apache Spark~\cite{CUSTOM:web/Spark} is a data processing framework that implements a DAG-based processing framework, provides an SQL optimizer, performs driver-based iterations, and treats stream computations as micro-batches. In contrast, Flink is the only system that 
\begin{inparaenum}[i)]
  \item supports an optimizer that can optimize DAG programs beyond SQL queries,
  \item is able to perform very efficient iterative processing natively,
  \item includes stream processing as a first-class citizen, enabling more complex use cases than micro-batches.
\end{inparaenum}

\para{Stream Processing.} There is a wealth of work on streaming systems, that includes commercial systems like StreamBase, Microsoft StreamInsight, and IBM InfoSphere Streams. Many of these systems are based on research in the database community in projects such as PSoup/Telegraph \cite{chandrasekaran2003psoup}, Aurora/Borealis~\cite{abadi2005design}, Stanford STREAM~\cite{arasu2004stream}, Trill~\cite{chandramouli2014trill} and IBM System S~\cite{gedik2008spade}. Most of the above systems are either
\begin{inparaenum}[i)]
  \item academic prototypes,
  \item closed-source commercial products, or 
  \item do not scale the computation horizontally on clusters of commodity servers.
\end{inparaenum}
Newer open source streaming systems that scale horizontally, such as Apache Storm~\cite{CUSTOM:web/Storm} and Apache Samza~\cite{CUSTOM:web/Samza}, provide low level APIs and offer only at-least-once and at-most-once guarantees. MillWheel~\cite{akidau2013millwheel}, a closed-source system at Google provides exactly-once guarantees with low latency and is used by Google Dataflow~\cite{akidau2015dataflow} to implement  out-of-order stream analytics. To the best of our knowledge, Flink is the only open-source project that
\begin{inparaenum}[i)]
  \item offers high level APIs and processing libraries,
  \item provides state management with exactly-once guarantees and
  \item achieves high throughput and low latency, serving both batch and true streaming computations equally well.
\end{inparaenum}

>>>>

\para{Batch Processing.} Apache Hadoop\footnote{https://hadoop.apache.org/} is the most popular open source system for large-scale data analysis that is based on the MapReduce paradigm~\cite{DBLP:journals/cacm/DeanG08}. Dryad~\cite{isard2007dryad}, introduced embedded user-defined functions in general DAG-based dataflows and was used by SCOPE~\cite{scopeOptimizer} which added a language and an SQL optimizer on top of it. Apache Tez\footnote{https://tez.apache.org/} can be seen as an open source implementation of the ideas proposed in Dryad. MPP databases \cite{dewitt1990gamma}, and recent open-source implementations~\cite{CUSTOM:web/Drill,kornacker2015impala} restrict their API to SQL variants. Similar to Flink, Apache Spark\footnote{spark.apache.org} is a data processing framework that implements a DAG-based processing framework, provides an SQL optimizer, performs driver-based iterations, and treats stream computations as micro-batches. In contrast, Flink is the only system that 
\begin{inparaenum}[i)]
  \item supports an optimizer that can optimize DAG programs beyond SQL queries,
  \item is able to perform very efficient iterative processing natively,
  \item includes stream processing as a first-class citizen, enabling more complex use cases than micro-batches.
\end{inparaenum}

\para{Stream Processing.} There is a wealth of work on streaming systems, that includes commercial systems like StreamBase, Microsoft StreamInsight, and IBM InfoSphere Streams. Many of these systems are based on research in the database community in projects such as PSoup/Telegraph \cite{chandrasekaran2003psoup}, Aurora/Borealis~\cite{abadi2005design}, Stanford STREAM~\cite{arasu2004stream}, Trill~\cite{chandramouli2014trill} and IBM System S~\cite{gedik2008spade}. Most of the above systems are either
\begin{inparaenum}[i)]
  \item academic prototypes,
  \item closed-source commercial products, or 
  \item do not scale the computation horizontally on clusters of commodity servers.
\end{inparaenum}
Newer open source streaming systems that scale horizontally, such as Apache Storm\footnote{http://storm.apache.org/} and Apache Samza\footnote{http://samza.apache.org/}, provide low level APIs and offer only at-least-once and at-most-once guarantees. MillWheel~\cite{akidau2013millwheel}, a closed-source system at Google provides exactly-once guarantees with low latency and is used by Google Dataflow~\cite{akidau2015dataflow} to implement  out-of-order stream analytics. To the best of our knowledge, Flink is the only open-source project that
\begin{inparaenum}[i)]
  \item offers high level APIs and processing libraries,
  \item provides state management with exactly-once guarantees and
  \item achieves high throughput and low latency, serving both batch and true streaming computations equally well.
\end{inparaenum}
